Computational methods in landscape ecology

### 2024-08-30 Manuscript.tex
Adding some references

Resistance surfaces are one of the most popular approaches to model
connectivity and developed from simple isolation-by-distance approaches.
The approach uses environmental layers and biodiversity data or expert
opinions to estimate the cost and of movement through landscapes (Zeller
et al. 2012). The latest developments of resistance surfaces have
recently been reviewed (Dutta et al. 2022). The resistance surface can
be used to model movement based on different approaches such as
leas-cost paths (Adriaensen et al. 2003; Etherington 2016) or circuit
theory (McRae et al. 2008). However, resistance surfaces are not free of
critique including missing spatial-temporal variability or context
dependency (Unnithan Kumar et al. 2022b; Unnithan Kumar and Cushman
2022). Thus, recent developments include the combination with more
processes-based modelling approaches (Schumaker and Brookes 2018; Bocedi
et al. 2021; Unnithan Kumar et al. 2022a), and modelling of dynamic
landscape connectivity (Zeller et al. 2020).

@@

\subsection{Various}\label{various}

@@

@@

@@

@@

>>>>

Resistance surfaces are one of the most popular approaches to model
connectivity and developed from simple isolation-by-distance approaches.
The approach uses environmental layers and biodiversity data or expert
opinions to estimate the cost and of movement through landscapes (Zeller
et al. 2012). The latest developments of resistance surfaces have
recently been reviewed (Dutta et al. 2022). The resistance surface can
be used to model movement based on different approaches such as
leas-cost paths (Adriaensen et al. 2003; Etherington 2016) or circuit
theory (McRae et al. 2008). However, resistance surfaces are not free of
critique including missing spatial-temporal variability or context
dependency (Unnithan Kumar et al. 2022b; Unnithan Kumar and Cushman
2022). Thus, recent developments include the combination with more
processes-based modelling approaches (Schumaker and Brookes 2018; Bocedi
et al. 2021; Unnithan Kumar et al. 2022a), modelling of dynamic
landscape connectivity (Zeller et al. 2020), inclusion of stochasticity
and spatial context (Ribeiro et al. 2017), using genetic optimization
algorithms (Peterman 2018), or computational improvements (Leonard et
al. 2017).

@@

Graph theory describes a landscape as a set of nodes representing
habitat patches and edges as potential links between the nodes. Thus,
graph theory is a powerful tool to model landscape connectivity,
especially in a metapopulation context (Urban and Keitt 2001).

\subsection{Other approaches}\label{other-approaches}

@@

\bibitem[\citeproctext]{ref-Leonard2017}
Leonard PB, Duffy EB, Baldwin RF, et al (2017) Glow: Software for
modelling circuit theory-based connectivity at any scale. Methods in
Ecology and Evolution 8:519--526.
\url{https://doi.org/10.1111/2041-210X.12689}

@@

\bibitem[\citeproctext]{ref-Peterman2018}
Peterman WE (2018) ResistanceGA: An r package for the optimization of
resistance surfaces using genetic algorithms. Methods in Ecology and
Evolution 9:1638--1647. \url{https://doi.org/10.1111/2041-210X.12984}

@@

\bibitem[\citeproctext]{ref-Ribeiro2017}
Ribeiro JW, Silveira Dos Santos J, Dodonov P, et al (2017) LandScape
corridors ( {\textsc{lscorridors}} ): A new software package for
modelling ecological corridors based on landscape patterns and species
requirements. Methods in Ecology and Evolution 8:1425--1432.
\url{https://doi.org/10.1111/2041-210X.12750}

@@

\bibitem[\citeproctext]{ref-Urban2001}
Urban D, Keitt T (2001) Landscape connectivity: A graph-theoretic
perspective. Ecology 82:1205--1218.
\url{https://doi.org/10.1890/0012-9658(2001)082\%5B1205:LCAGTP\%5D2.0.CO;2}

### 2024-08-30 Sections/03-connectivity.qmd
Adding some references

<!-- 
Connectivity modelling in conservation science: https://doi.org/10.1038/s41598-022-20370-w
-->

@@

Resistance surfaces are one of the most popular approaches to model connectivity and developed from simple isolation-by-distance approaches. 
The approach uses environmental layers and biodiversity data or expert opinions to estimate the cost and of movement through landscapes [@Zeller2012].
The latest developments of resistance surfaces have recently been reviewed [@Dutta2022].
The resistance surface can be used to model movement based on different approaches such as leas-cost paths [@Adriaensen2003; @Etherington2016] or circuit theory [@McRae2008].
However, resistance surfaces are not free of critique including missing spatial-temporal variability or context dependency [@UnnithanKumar2022; @UnnithanKumar2022a].
Thus, recent developments include the combination with more processes-based modelling approaches [@Schumaker2018; @Bocedi2021; @UnnithanKumar2022b], and modelling of dynamic landscape connectivity [@Zeller2020]. <!-- MH: Not complete yet -->

<!--
lscorridors: https://doi.org/10.1111/2041-210X.12750
Complementary landscape genetics and habitat modelling approaches: https://doi.org/10.1007/s10980-019-00789-9
https://doi.org/10.1111/2041-210X.12689
https://doi.org/10.1016/j.ecolind.2016.03.032
-->

@@

<!--
https://doi.org/10.1007/s10980-020-00998-7
https://doi.org/10.1007/s10980-020-01164-9
-->

## Various

<!--
https://doi.org/10.1890/14-1690.1
https://doi.org/10.1016/j.ecolind.2016.03.032
https://doi.org/10.1007/s10980-019-00935-3
https://doi.org/10.1126/science.abg1780
-->

>>>>

<!-- 
Connectivity modelling in conservation science https://doi.org/10.1038/s41598-022-20370-w
-->

@@

Resistance surfaces are one of the most popular approaches to model connectivity and developed from simple isolation-by-distance approaches. 
The approach uses environmental layers and biodiversity data or expert opinions to estimate the cost and of movement through landscapes [@Zeller2012].
The latest developments of resistance surfaces have recently been reviewed [@Dutta2022].
The resistance surface can be used to model movement based on different approaches such as leas-cost paths [@Adriaensen2003; @Etherington2016] or circuit theory [@McRae2008].
However, resistance surfaces are not free of critique including missing spatial-temporal variability or context dependency [@UnnithanKumar2022; @UnnithanKumar2022a].
Thus, recent developments include the combination with more processes-based modelling approaches [@Schumaker2018; @Bocedi2021; @UnnithanKumar2022b], modelling of dynamic landscape connectivity [@Zeller2020], inclusion of stochasticity and spatial context [@Ribeiro2017], using genetic optimization algorithms [@Peterman2018], or computational improvements [@Leonard2017].

<!-- What about multi-species approaches? -->

@@

Graph theory describes a landscape as a set of nodes representing habitat patches and edges as potential links between the nodes.
Thus, graph theory is a powerful tool to model landscape connectivity, especially in a metapopulation context [@Urban2001].

<!--
Coupling landscape graph modeling and biological data https://doi.org/10.1007/s10980-020-00998-7
Temporal dynamics of landscape connectivity in ever-changing landscapes: a literature review https://doi.org/10.1007/s10980-021-01277-9
Some methodological considerations to improve landscape planning https://doi.org/10.1016/j.ecolind.2017.05.052
Comparison of landscape graph modelling methods https://doi.org/10.1007/s10980-020-01164-9

-->

## Other approaches

<!--
A stochastic movement simulator improves estimates of landscape connectivity https://doi.org/10.1890/14-1690.1
Landscape connectivity modeling from the perspective of animal dispersal  https://doi.org/10.1007/s10980-019-00935-3
Big-data approaches https://doi.org/10.1126/science.abg1780
-->
